### 2025-09-06 cognitive-bias.qmd
20250906 - page numbers

*Overestimation* involves believing that one will perform better than one actually performs.
It thus reflects a greater mean confidence than overall accuracy [@Kahneman1996].
In general, people tend to show excessive confidence in their beliefs; for instance, when people express 90% confidence about their belief, they may be correct only about 75% of the time [@Kahneman1996].
As an example of overconfidence in football, early picks in the NFL draft tend to be considerably overvalued because teams believe their predictions will be more accurate than they actually are [@Massey2013].
The probability that a player is better than the next player chosen in the draft at the same position is ~52%, or basically a coin flip [@Massey2013].
As noted by @Massey2013 (p. ) "...decision makers often know less than they think they know...The problem is not that future performance is difficult to predict, but that decision makers do not appreciate how difficult it is."
This suggests that a) the presence of high stakes for the prediction does not eliminate overconfidence, and that b) teams should "trade down" to acquire more, later picks if they possess very early picks [@Massey2013].
Overestimation can be identified with a [calibration plot](#sec-calibrationPlot) of the predicted performance versus actual performance, where the person's predicted performance is systematically higher (in at least some cases) than their actual performance.
Overestimation corresponds to the "[overprediction](#sec-calibration)" form of [miscalibration](#sec-calibration).

>>>>

*Overestimation* involves believing that one will perform better than one actually performs.
It thus reflects a greater mean confidence than overall accuracy [@Kahneman1996].
In general, people tend to show excessive confidence in their beliefs; for instance, when people express 90% confidence about their belief, they may be correct only about 75% of the time [@Kahneman1996].
As an example of overconfidence in football, early picks in the NFL draft tend to be considerably overvalued because teams believe their predictions will be more accurate than they actually are [@Massey2013].
The probability that a player is better than the next player chosen in the draft at the same position is ~52%, or basically a coin flip [@Massey2013].
As noted by @Massey2013 (pp. 1493–1494), "The task of picking players...is an extremely difficult one...decision makers often know less than they think they know...The problem is not that future performance is difficult to predict, but that decision makers do not appreciate how difficult it is."
This suggests that a) the presence of high stakes for the prediction does not eliminate overconfidence, and that b) teams should "trade down" to acquire more, later picks if they possess very early picks [@Massey2013].
Overestimation can be identified with a [calibration plot](#sec-calibrationPlot) of the predicted performance versus actual performance, where the person's predicted performance is systematically higher (in at least some cases) than their actual performance.
Overestimation corresponds to the "[overprediction](#sec-calibration)" form of [miscalibration](#sec-calibration).
